---
title: "Final Project"
author: "Eddie Perez"
date: "1/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

## R Markdown

This is a final project to show off what you have learned. Select your data set from the list below: 
http://vincentarelbundock.github.io/Rdatasets/ (click on the csv index for a list). Another good source is
found here:
https://https://archive.ics.uci.edu/ml/datasets.html
The presentation approach is up to you but it should contain the following:
1. Data Exploration: This should include summary statistics, means, medians, quartiles, or any 
other relevant information about the data set. Please include some conclusions in the R 
Markdown text.
2. Data wrangling: Please perform some basic transformations. They will need to make sense but 
could include column renaming, creating a subset of the data, replacing values, or creating new 
columns with derived data (for example – if it makes sense you could sum two columns 
together)
3. Graphics: Please make sure to display at least one scatter plot, box plot and histogram. Don’t 
be limited to this. Please explore the many other options in R packages such as ggplot2.
4. Meaningful question for analysis: Please state at the beginning a meaningful question for 
analysis. Use the first three steps and anything else that would be helpful to answer the 
question you are posing from the data set you chose. Please write a brief conclusion paragraph
in R markdown at the end.
5. BONUS – place the original .csv in a github file and have R read from the link. This will be a very 
useful skill as you progress in your data science education and career.

## Analysis

I decided to analyze who or what population of people servied in 1985 smoked the most, Based on gender marital status employment, etc..

```{r}
# Load the data from my github part of the bonus same file as local 
file <- "https://raw.githubusercontent.com/Eperez54/R-Final-Project/main/CPS1985.csv?token=GHSAT0AAAAAABQPZB67AVPVCY2UGTUPINNMYPN54CA"

smokerdf <- read.csv(file, TRUE, ",") 
smokerdf <- data.frame(smokerdf) %>% na.omit() # not counting N/A Values

head(smokerdf)
```

## Data Exploration
```{r}
# Print summary
summary(smokerdf)
```

## Let's See the Column names
```{r}
colnames(smokerdf,TRUE, col)
```

```{r}
#print Means and Medians
means <- sapply(smokerdf[, c("wage", "education", "experience", "age", "ethnicity", "region", "gender",
                             "occupation", "sector", "union", "married")], mean)
medians <- sapply(smokerdf[, c("wage", "education", "experience", "age", "ethnicity", "region", "gender",
                             "occupation", "sector", "union", "married")], median)
mean_mediandf <- data.frame(means, medians)
mean_mediandf
```

## Create a new data frame with a subset of main data frame
```{r}
smoker_subset <- smokerdf[1:40,c("wage", "education", "experience", "age", "ethnicity", "region", "gender",
                             "occupation", "sector", "union", "married")]
head(smoker_subset)
```

```{r}
tail(smoker_subset)
```

```{r}
# Rename some columns in my subset data
colnames(smoker_subset) <- c("Salary", "Education", "Experience", "Age", "Race", "Region", "Sex",
                             "Occupation", "Sector", "union", "Married")
colnames(smoker_subset)
```

```{r}
head(smoker_subset)
```

```{r}
tail(smoker_subset)
```

## I want to see summary of my new data set

```{r}
summary(smoker_subset)
```

```{r}
#print Means and Medians
means2 <- sapply(smoker_subset[, c("Salary", "Education", "Experience", "Age", "Race", "Region", "Sex",
                             "Occupation", "Sector", "union", "Married")], mean)
medians2 <- sapply(smoker_subset[, c("Salary", "Education", "Experience", "Age", "Race", "Region", "Sex",
                             "Occupation", "Sector", "union", "Married")], median)
mean_mediandf2 <- data.frame(means2, medians2)
mean_mediandf2
```
## Lets see some graphs 

```{r}
summary(smoker_subset$Age)
```

```{r}
hist(smoker_subset$Age,main = "Age Histogram", xlab = "Age")
```
## Observation
Most people that smoke are in their late 20's and early 30's
lets see on their education level. lets see this information in a bar graph

```{r}
summary(smoker_subset$Education)
```
```{r}
summary(smoker_subset$Salary)
```
```{r}
tn <- table(smoker_subset$Education)

barplot(tn, main= "Smoker Based Education", xlab="Highest Education Finished", 
        col=c("yellow","orange", "blue","green","red"),
        legend= rownames(tn), beside = TRUE)
tn

```
## Observation
the most smokers have no college education. only finished HS.


```{r}
summary(smoker_subset$Education)
```

```{r}
summary(smoker_subset$Salary)
```

```{r}
Hist <- ggplot(smoker_subset,aes(x=Education)) + 
  geom_histogram(fill = "gray",bins=10)

Hist
```
I like how it look in a barplot

```{r}
Hist <- ggplot(smoker_subset,aes(x=Age)) + 
  geom_histogram(fill = "gray",bins=10)

Hist
```

```{r}
Hist <- ggplot(smoker_subset,aes(x=Experience)) + 
  geom_histogram(fill = "gray",bins=10)

Hist
```


## Interesting
Most smokers have over 10 year of job experience.

The scatter plot belowe will show the relationship between years of experience and Education. Many of the smoker have graduated from High School.

```{r}
#Scatter Plot 
ScatterPlot <- ggplot(smoker_subset,aes(x=Experience, y=Age, color = factor(Age))) + geom_point(size=2.5)
ScatterPlot
```



The more experience you have the more likely you will be a smoker?

```{r}
ScatterPlot <- ggplot(smoker_subset,aes(x=Experience, y=Education, color = factor(Education))) + geom_point(size=2.5)
ScatterPlot
```
## It is pretty constant that most High school graduate in the late 80's where smokers


```{r}
ScatterPlot <- ggplot(smoker_subset,aes(x=Race, y=Salary, color = factor(Race))) + geom_point(size=2.5)
ScatterPlot
```

## Conclusion

#### while wrangling and exploration i came to a realization that the more experience and years you have in a given job you where a smoker. maybe the job had too much pressure or maybe there was no campaing to help inform of the harm of smoking. It was very alarming that most smoker where High school only graduate with no college experience. I would love to verify this information with more current data, to see if the FDA and the surgeon general have had any impact on lowering these number. Furthermore most smokers are mainly Caucasians.